Prediction of 90-day mortality after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden, England and Wales

### 2020-09-03 reports/figs.Rmd
response from NH and EB

(ref:br1) ~~Outer steps of the bootstrap ranking procedure. All included patients (A) were re-sampled with replacements 100 times (B). For each bootstrap sample, the inner process (C; as depicted in figure \@ref(fig:br2)) was applied. The final output was a table with all potential predictors ranked by the number of times they got selected (D). Potential predictors selected at least once were included in the main model, and each predictor selected at least 33 out of the 100 times, were included in a reduced model as well.~~

@@

(ref:br2) ~~Inner steps of the bootstrap ranking procedure. For each bootstrap sample (1), 100 new bootstrap samples were created (2). Logistic LASSO-regression was applied within each sample (3). Absolute values of the estimated coefficient values (whereof some shrunk to 0 by the LASSO), were averaged as a measure of variable importance (4). Variables with their estimated variable importance above an estimated break-point from linear piece-wise regression (5) were kept as potential predictors (6).~~

@@

```{r rocvalidation, fig.cap = "Receiver Operation Characteristics (ROC) curves. There was no difference when applying the reduced model to the external validation cohort from England and Wales ,compared to the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/roc_validation.eps")
knitr::include_graphics("../graphs/images.png")
```

>>>>

(ref:br1) ~~Outer steps of the bootstrap ranking procedure. All included patients (A) were re-sampled with replacements 100 times (B). For each bootstrap sample, the inner process (C; as depicted in figure \@ref(fig:br2)) was applied. The final output was a table with all potential predictors ranked by the number of times they got selected (D). Potential predictors selected at least once were included in the main model, and each predictor selected at least 33 out of the 100 times, were included in a reduced model as well.[@Bulow2020a]~~

@@

(ref:br2) ~~Inner steps of the bootstrap ranking procedure. For each bootstrap sample (1), 100 new bootstrap samples were created (2). Logistic LASSO-regression was applied within each sample (3). Absolute values of the estimated coefficient values (whereof some shrunk to 0 by the LASSO), were averaged as a measure of variable importance (4). Variables with their estimated variable importance above an estimated break-point from linear piece-wise regression (5) were kept as potential predictors (6).[@Bulow2020a]~~

@@

```{r rocvalidation, fig.cap = "Receiver Operation Characteristics (ROC) curves. There was no difference when applying the reduced model to the external validation cohort from England and Wales, compared to the Swedish derivation cohort."}
# knitr::include_graphics("../graphs/roc_validation.eps")
knitr::include_graphics("../graphs/images.png")
```

### 2020-09-03 reports/tabs.Rmd
response from NH and EB

(ref:comorbidity) Categorization of comorbidities based on diagnostic groups according to the Charlson and Elixhauser comorbidity indices. ~~Each sub-condition was firsdt classified based on data from the Swedish National patient register and from the Hospital Episodes statistics registry in England and Wales, coded by the International classification of diseases veriosn 10 (ICD-10), with categorization defined in table 1 and 2 from Quan et al.[Quan2005]~~. (CNS = central nervous system.)

```{r tabcategorization}
tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`)
  ) %>% 
  knitr::kable(
    caption = "(ref:comorbidity)")

@@

```{r tab1}
table1_combined_tot %>% 
  mutate(what = gsub("Heart infarct", "Myocardial infarction", what)) %>% 
knitr::kable(
    caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anesthesiologists classification. CNS = central nervous system.")
```

@@

```{r brlprop}
t <- 
  brlasso_tbl_selected %>% 
  mutate(
    variable = clean_names(variable),
    variable = gsub("= 2", "II", variable),
    variable = gsub("= 3", "III", variable),
    variable = gsub("Heart infarct", "Myocardial infarction", variable)
  )

@@

```{r brlreduced}
brlasso_tbl_coefs %>% 
  select(-math) %>%
  mutate(
    term = gsub("= 2", "II", term),
    term = gsub("= 3", "III", term)
  ) %>% 
knitr::kable(
  caption = 'Estimated coefficients (beta) and odds ratios (OR) with 95 % confidence intervals for the reduced model.'
)
```

>>>>

(ref:comorbidity) Categorization of comorbidities based on diagnostic groups according to the Charlson and Elixhauser comorbidity indices. ~~Each sub-condition was first classified based on data from the Swedish National patient register and from the Hospital Episodes statistics registry in England and Wales, coded by the International classification of diseases version 10 (ICD-10), with categorization defined in table 1 and 2 from Quan et al.[@Quan2005]~~ (CNS = central nervous system.)

```{r tabcategorization}
tab_categorization %>%
  ungroup() %>% 
  mutate(
    `Comorbidities by groups` = 
      gsub("Heart infarct", "Myocardial infarction", `Comorbidities by groups`),
    `Comorbidities by groups` = 
      gsub("Obesity", "~~Diagnosed~~ obesity", `Comorbidities by groups`)
  ) %>% 
  mutate_all(~ gsub("nemia", "naemia", .)) %>% 
  mutate_all(~ gsub("tumor", "tumour", .)) %>% 
  knitr::kable(
    caption = "(ref:comorbidity)")

@@

@@

```{r tab1}
table1_combined_tot %>% 
  mutate(
    what = gsub("Heart infarct", "Myocardial infarction", what),
    what = gsub("Obesity", "~~Diagnosed~~ obesity", what),
    what = gsub("Anemia", "Anaemia", what)
  ) %>% 
knitr::kable(
    caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anaesthesiologists classification. CNS = central nervous system.")
```

```{r brlprop}
t <- 
  brlasso_tbl_selected %>% 
  mutate(
    variable = clean_names(variable),
    variable = gsub("= 2", "II", variable),
    variable = gsub("= 3", "III", variable),
    variable = gsub("Heart infarct", "Myocardial infarction", variable),
    variable = gsub("Obesity", "~~Diagnosed~~ obesity", variable),
    variable = gsub("Anemia", "Anaemia", variable)
  )

@@

```{r brlreduced}
brlasso_tbl_coefs %>% 
  select(-math) %>%
  mutate(
    term = gsub("= 2", "II", term),
    term = gsub("= 3", "III", term),
    term = gsub("Obesity", "~~Diagnosed~~ obesity", term)
  ) %>% 
knitr::kable(
  caption = 'Estimated coefficients (beta) and odds ratios (OR) with 95 % confidence intervals for the reduced model.'
)
```
